### 2010-03-25 docs/architecture/architecture.tex
add a description on the GNUmakefile

>>>>

Different problems in MAESTRO are contained in individual
sub-directories under {\tt MAESTRO/}.  The minimum requirement to
define a new problem would be a {\tt GNUmakefile} which describes how
to build the application and an input file which lists the runtime
parameters.  The problem-specific executable is built in the problem
directory by typing {\tt make}.  Source files are found automatically
by searching the directories listed in the {\tt GNUmakefile}.

\subsection{The {\tt GNUmakefile}}

A basic {\tt GNUmakefile} begins with:
\begin{verbatim}
  NDEBUG := t
  MPI    :=
  OMP    :=
\end{verbatim}
Here, {\tt NDEBUG} is true is we are building an optimized executable.  Otherwise,
the debug version is built---this typically uses less optimization and adds various
runtime checks through compiler flags.  {\tt MPI} and {\tt OMP} are set to true if
we want to use either MPI or OpenMP for parallelization.

The MAESTRO build system knows what options to use for various compiler families.
The {\tt COMP} flag specifies which compiler to us.  Allowed values include
{\tt Intel}, {\tt gfortran}, {\tt PGI}, and {\tt PathScale}.  The specific details
of these choices are defined in the {\tt MAESTRO/mk/} directory.
\begin{verbatim}
  COMP := Intel
  MKVERBOSE := t
\end{verbatim}
{\tt MKVERBOSE} set to true will echo the build commands to the terminal as the
are executed.

The next set of lines tell the build system where to look for source
files.  A MAESTRO application is built from several pacakges (the
multigrid solver, an EOS, a reaction network, etc.).  
\begin{verbatim}
  FPARALLEL := ../..

  include $(FPARALLEL)/mk/GMakedefs.mak

  Fmdirs := boxlib \
            mg \
            MAESTRO \
            extern/EOS/helmeos \
            extern/EOS/conducteos \
            extern/constants \
            extern/networks/ignition \
            extern/VODE

  Fmincludes := extern/EOS/helmeos
\end{verbatim}
The {\tt FPARALLEL} variable should point to the top-level directory
of the source tree (i.e.\ {\tt fParallel/}).  Next we include the
compiler definitions, {\tt mk/GMakedefs.mak}.  The {\tt Fmdirs}
variable lists the individual directories in the {\tt fParallel/}
source tree that are used to build the problem.  Generally, one does
not need to include the problem directory itself, unless there are
unique source files found there, described in a {\tt GPackage.mak}
file.  The build system will attempt to build all of the files listed
in the various {\tt GPackage.mak} files found in the {\tt Fmdirs}
directories.  Furthermore, {\tt Fmdirs} will be will be added to the
{\tt make} {\tt VPATH}, which is the list of directories to search for
source files.  The problem directory will always be put first in the
{\tt VPATH}, so any source files placed there override those with the
same name found elsewhere in the source tree.  The {\tt Fmincludes} variable
lists all those directories that contain include files that are inserted into
the Fortran source at compile time via the {\tt include} statement.  Presently,
the only instance of this is with the Helmholtz general equation of state
found in {\tt extern/EOS/helmeos/}.

Runtime parameters listed in the {\tt MAESTRO/\_parameters} file are
parsed at compile time and the file {\tt probin.f90} is written and 
compiled.  This is a Fortran module that holds the values of the runtime
parameters and makes them available to any routine.  The variable
{\tt PROBIN\_PARAMETERS} should list all of the files that contain
runtime parameters for the current problem.  You must always
include the main {\tt ../\_parameters} file (i.e.\ {\tt MAESTRO/\_parameters}).
Sometimes a problem defines their own runtime parameters, in which
case, its parameter file should also appear in {\tt PROBIN\_PARAMETERS}.
\begin{verbatim}
  PROBIN_PARAMETERS := ../_parameters
\end{verbatim}

The final lines in the {\tt GNUmakefile} include the rules to actually
build the executable.
\begin{verbatim}
  include $(FPARALLEL)/MAESTRO/GMaestro.mak
  include $(FPARALLEL)/mk/GMakerules.mak
\end{verbatim}

As mentioned above, any source files placed in the problem directory
override a files with the same name found elsewhere in the source
tree.  This allows you to create a problem-specific version of any
routine.  Source files that are unique to this problem (i.e.\ there is
no file with the same name elsewhere in the source tree) need to be
listed in a file {\tt GPackage.mak} in the problem directory.
